---
output: github_document
---

<!-- badges: start -->
[![R-CMD-check](https://github.com/joaobtj/irritool/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/joaobtj/irritool/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->
  
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# irritool

The **irritool** package provides computational tools in R for irrigation.

## Installation

You can install the development version of irritool from [GitHub](https://github.com/) with:

``` {r eval=FALSE}
# install.packages("devtools")
devtools::install_github("joaobtj/irritool")
```

## Documentation and Tutorials

Detailed examples, including the complete workflow for calculating crop water requirements and simulating the daily soil water balance, are available in the package vignettes.

<<<<<<< HEAD
If you are using the package website, you can read the full tutorials in the Articles section:
https://joaobtj.github.io/irritool/
=======
## Example Usage

Below is a basic workflow showing how the package functions integrate to simulate a crop cycle.

```{r example, warning=FALSE, message=FALSE}
library(irritool)

# 1. Obtain climate data for a coordinate (requires local NetCDF files)
# As this is a reproducible example, we leave the extraction commented out:
# climate_data <- extract_brdwgd_point(
#   target_longitude = -50.59,
#   target_latitude = -27.28,
#   weather_variables = c("pr", "ETo"),
#   nc_files_directory = "path/to/files"
# )

# For this example, we will use mock climate data:
cycle_days <- 65
set.seed(123)
rain_data <- sample(c(rep(0, 60), runif(5, 5, 25)))
eto_data <- runif(cycle_days, 3, 5)

# 2. Generate the Kc curve for the crop (e.g., Lettuce)
lettuce_kc <- calc_kc_curve(
  kc_points = c(0.7, 1.0, 0.95),
  stage_lengths = c(15, 20, 20, 10),
  crop = "Lettuce"
)

# View the generated plot
print(lettuce_kc$kc_plot)
```

```{r example_balance}
# 3. Simulate the soil water balance
water_balance <- calc_water_balance(
  et0 = eto_data,
  rainfall = rain_data,
  daily_kc_values = lettuce_kc$kc_serie, # Consuming the output from the previous function
  root_depth = 300,
  theta_fc = 0.30,
  theta_wp = 0.15,
  depletion_factor = 0.55,
  irrigation_rule = "threshold"
)

# View the first few days of the soil water balance
head(water_balance$water_balance_data)

# View the summary depths (totals)
print(water_balance$summary_depths)
```

>>>>>>> bb4199070b43fa16aa226864cc6e4f656aeb0053
